Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction

### 2014-09-29 article.Rmd
Nuanced some langauge
Plot tweak, citations

```{r plot-names-that-change}
plot_names <-  c("madison",
                 "keanu",
                 "addison",
                 "jan",
                 "sydney",
                 "morgan",
                 "leslie",
                 "kendall")

@@

One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013]. Our own analysis reveals a more complicated reality. The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors. This disparity is put into even sharper relief when set against the relative parity achieved by women in producing history dissertations. Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession. Our program allows us to quantitatively measure the scope of the problem on a large scale and how it has changed over time.

@@

@@

[Genderize.io 2014-] *Genderize.io*. Accessed 15 September 2014, http://genderize.io/.

@@

[Healy 2013] Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/

@@

[Lunbeck 2005] Lunbeck, Elizabeth, et al. The Status of Women in the Historical Profession. American Historical Association Committee on Women Historians (2005): http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf.

[Muenchen 2014-] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.

@@

[Piatetsky 2013] Piatetsky, Gregory. "Top Languages for analytics, data mining, data science," *KDN Nuggets*. 27 August 2013, http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html.

@@

[Social Security Administration 2014-a] Social Security Administration. "Baby Names from Social Security Card Applications-National Level Data". Data.gov dataset ID US-GOV-SSA-338. Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data

[Social Security Administration 2014-b] Social Security Administration. "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339. Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-

@@

[Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): http://www.jstatsoft.org/v59/i10.

@@

[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.

>>>>

```{r plot-names-that-change}
plot_names <-  c("madison",
                 #"keanu", #keanu doesn't seem to actually flip, or at least doesn't show up on the chart?
                 "addison",
                 "jan",
                 "sydney",
                 "morgan",
                 "leslie",
                 "kendall")

@@

One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013]. Our own analysis reveals that the story is different for book publishing than for employment. The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors. This disparity is put into even sharper relief when set against the relative parity achieved by women in producing history dissertations. Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession. Our program allows us to quantitatively measure the scope of the problem on a large scale and how it has changed over time.

@@

[American Historical Review n.d.] American Historical Review. "Book Reviewing in the AHR." http://www.historians.org/publications-and-directories/american-historical-review/book-review-guide. Accessed 29 September 2014.

@@

[Genderize.io n.d.] *Genderize.io*. http://genderize.io/. Accessed 15 September 2014.

@@

[Healy 2013] Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, <http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/>.

@@

[Lunbeck 2005] Lunbeck, Elizabeth, et al. "The Status of Women in the Historical Profession." American Historical Association Committee on Women Historians. May 2005, <http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf>.

[Muenchen n.d.] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. <http://r4stats.com/articles/popularity/>. Accessed 12 September 2014.

@@

[Piatetsky 2013] Piatetsky, Gregory. "Top Languages for analytics, data mining, data science," *KDN Nuggets*. 27 August 2013, <http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html>.

@@

[Social Security Administration 2014-a] Social Security Administration. "Baby Names from Social Security Card Applications-National Level Data". Data.gov dataset ID US-GOV-SSA-338. Accessed 31 August 2014, <http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data>.

[Social Security Administration 2014-b] Social Security Administration. "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339. Accessed 31 August 2014, <http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of->.

@@

[Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): <http://www.jstatsoft.org/v59/i10>.

@@

[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. <http://data.worldbank.org/indicator/SP.DYN.LE00.IN>. Accessed 12 September 2014
